---
title: "Homework 2, Biostatistics"
output:
  html_document: default
  pdf_document: default
---

```{r,message=F,warning=F}
library(gmodels)
library(epiR)
library(exact2x2)
library(fmsb)
library(psych)
library(BSDA)
library(MASS)
library(DiffXTables)
```
# Q1
```{r}
Q1_data=array(c(8,52,5,164,25,29,21,138,50,27,61,208),c(2,2,3),dimnames =list(c('Case','Control'),c('H','L'),c('35~44','45~54','55~64')))
Q1_data
```

## Q1 (a)
```{r}
mantelhaen.test(Q1_data, correct=F)
```
from Mantel-Haenszel test , we can see that p-value < 2.2e-16 , so we reject $H_0$(i.e odds ratio is not equal to 1) , and the common odds ratio is 5.882095

## Q1 (b)
compare 4 tables (group by age and crude) of odds ratio
```{r}
epi.2by2(Q1_data[,,1], method="case.control")
epi.2by2(Q1_data[,,2], method="case.control")
epi.2by2(Q1_data[,,3], method="case.control")
epi.2by2(Q1_data[,,1]+Q1_data[,,2]+Q1_data[,,3], method="case.control")
```
find that 4 odds ratio are difference($odds\ ratio_1 =5.05$,$odds\ ratio_2 =5.67$,$odds\ ratio_3 =6.31$,$odds\ ratio_{crude} =4.51$) , so age is confounding and stratify by age is reasonable , if we not stratify by age the analysis have Simpson's Paradox

# --------------------------------------------------------------------------------

# Q2
```{r}
Q2_data = array(c(40,16,20,3),c(2,2),dimnames = list(c('A_success','A_failure'),c('B_success','B_success')))
Q2_data
```

matched data , McNemar's (Chi-squared) test for antibiotics  vs. gonorrhea and Exact McNemar's test  for antibiotics  vs. gonorrhea
```{r}
mcnemar.test(Q2_data, correct=F)
mcnemar.exact(Q2_data)
```
whether approximate or exact p-value > 0.05,do not reject $H_0$(i.e antibiotics A and B in treating gonorrhea effectiveness  is not different) ,  

# --------------------------------------------------------------------------------

# Q3
```{r}
Q3_data=array(c(35,21,13,249),c(2,2),dimnames = list(c('B_reviewer_+','B_reviewer_-'),c('A_reviewer_+','A_reviewer_-')))
Q3_data
```
Measure of agreement and its test
```{r}
Kappa.test(Q3_data)
```
since p-value < 0.05 , reject $H_0$(i.e Kappa coefficient not equal to 0), and estimates is 0.6096187 , mean is Substantial agreement

# --------------------------------------------------------------------------------

# Q4

## Q4(a)
$X_1,...,X_{10}$ ~ $Bin(20,p)$ where p is Pr($\theta\ \notin CI$) = 0.1
so Pr($X_1 \leq 3$) = $C^{20}_0 0.9^{20} + C^{20}_1 0.9^{19}0.1^1 + C^{20}_2 0.9^{18}0.1^2 + C^{20}_3 0.9^{17}0.1^3$

so each students  have no more than 3 confidence intervals out of 20 is $(Pr(X_1 \leq 3))^{10}$

```{r}
probability_no_more_than_3 = pbinom(3,20,0.1)
probability_no_more_than_3^10
```
probability that all 10 students each have no more than 3 confidence intervals that “miss” (i.e., do not contain the true parameter value) out of 20 is 0.242118

## Q4(b)
we can use goodness of fit test , $H_0 : X\sim Bin(20,0.1)$ vs.$H_1: not\ H_0$

and since the number of X=0 and X=1 is <5 so combin then and the number of X=4 and X>5 is <5 similar
```{r}
Q4_data=c(2+4,12,77,5+0)
chisq.test(Q4_data, p = c(dbinom(0,20,0.1)+dbinom(1,20,0.1),dbinom(2,20,0.1),dbinom(3,20,0.1),dbinom(4,20,0.1)+1-pbinom(4,20,0.1)))
```
p-value = 2.2e-16 < 0.05 , reject $H_0$

# --------------------------------------------------------------------------------

# Q5
$H_0: median\ of\ case = median\ of\ control$ vs. $H_0: median\ of\ case \neq median\ of\ control$

nonparametric test , so use Wilcoxon rank-sum test
```{r}
Q5_data_1 = c(60, 25, 6, 8, 5, 10, 25, 15, 10)
Q5_data_2 = c(13, 20, 15, 7, 75, 120, 10, 100, 9, 25, 30)
# exact test
wilcox.test(Q5_data_1, Q5_data_2, paired=F, exact=T, alternative="two.side")
# Normal approximation
wilcox.test(Q5_data_1, Q5_data_2, paired=F, exact=F, correct=F, alternative="two.side")
```
whether exact or approximation p-value > 0.05 , so do not reject $H_0$ (i.e have equal median) 


# --------------------------------------------------------------------------------

# Q6
```{r}
Q6_data = matrix(c(243,16,83,36,60,32,31,31,39,12),c(2,5),dimnames = list(c('Number_children_with_caries','Number_children_with_caries_free_teeth'),c('Essex','Slough','Harwick','Burnham','Meresa')))
```

```{r}
Q6_data
```

```{r}
two_cities_names = c()
p_value=c()
for (i in c(1:4)){
  for (j in c((i+1):5)){
  two_cities_names = c(two_cities_names,paste(colnames(Q6_data)[i],      colnames(Q6_data)[j], sep="_"))
  m = matrix( c(Q6_data[,i], Q6_data[,j]),c(2,2))
  p_value = c(p_value,chisq.test(m)$p.value)
  }
}
names(p_value) = two_cities_names
p_value
```

# Bonferroni 
```{r}
alpha = 0.05
adjust_p_palue = p.adjust(p_value, method="bonferroni")
sig = adjust_p_palue < alpha
print(adjust_p_palue[sig])
```

# Sidak
```{r}
alpha = 0.05
adjust_p_palue = 1-(1-p_value)^(length(p_value))
sig = adjust_p_palue < alpha
print(adjust_p_palue[sig])
```

# Holm
```{r}
alpha = 0.05
adjust_p_palue = p.adjust(p_value, method="holm")
sig = adjust_p_palue < alpha
print(adjust_p_palue[sig])
```

# Hochberg
```{r}
alpha = 0.05
adjust_p_palue = p.adjust(p_value, method="hochberg")
sig = adjust_p_palue < alpha
print(adjust_p_palue[sig])
```

# Benjamini-Hochberg
```{r}
alpha = 0.05
adjust_p_palue = p.adjust(p_value, method="BH")
sig = adjust_p_palue < alpha
print(adjust_p_palue[sig])
```
We then plot ordered unadjusted p-values vs. indices compared to various multiple comparison procedures.
```{r}
len = length(p_value)
plot(c(1:len),sort(p_value), xlab="Indices", ylab="Unadjsuted p-values")
lines(1:len, (1:len)*alpha/len, col="red", lty=1)
text(len, len*alpha/len, labels="BH", col="red")
lines(1:len, rep(alpha/len, len), col="green", lty=1)
text(len, alpha/len, labels="Bonferroni", col="green")
lines(1:len, rep(1-(1-alpha)^{1/len}, len), col="blue", lty=2)
text(len, 1-(1-alpha)^{1/len}, labels="Sidak", col="blue")
lines(1:len, alpha/(len:(len-(len-1))), col="brown", lty=2)
text(len, alpha/(len-(len-1)), labels="Holm/Hochberg", col="brown")
```

